---
title: "MRF [original]"
output: html_notebook
---



```{r}
setwd("/Volumes/Lesley_Chapman/American_University/Research_Assistantship/Code")
BALT = read.csv("./data/Baltimore_911_Calls_for_Service.csv")



# BALT = BALT[1:999999,]            # Only for coding speed

# Keep  "callDateTime" "priority" "location"       
# Erase "district" "description" "callNumber" "incidentLocation"
           
BALT = BALT[ , (names(BALT) %in% c("callDateTime", "priority", "location"))]

Ntotal = dim(BALT)[1]           # Ntotal = 2831903 calls

B = na.omit(BALT)

#B$month = as.numeric(substr(B$callDateTime,start=1,stop=2))
#B$day   = as.numeric(substr(B$callDateTime,start=4,stop=5))
#B$year  = as.numeric(substr(B$callDateTime,start=7,stop=10))
#B$hour  = as.numeric(substr(B$callDateTime,start=12,stop=13))
#B$min   = as.numeric(substr(B$callDateTime,start=15,stop=16))
#B$sec   = as.numeric(substr(B$callDateTime,start=18,stop=19))			# But sec = 0 always
#B$ampm  = substr(B$callDateTime,start=21,stop=22)
#B$hour = B$hour*(B$hour < 12) + 12*(B$hour < 12 & B$ampm=="PM") + 12*(B$hour == 12 & B$ampm=="PM")
#B$time = 60*B$min + 3600*B$hour

B$Day = as.Date(B$callDateTime,format="%m/%d/%Y")
B$t = as.numeric(B$Day)

# Check for any gaps in timeline
for (day in min(B$t):max(B$t)){ 
   if (sum(B$t==day)<10){ print(paste("Gap in timeline!!! Only",
         sum(B$t==day),"entries on day",day)) } 
                              }
 
# LATITUDE and LONGITUDE - from B$location of type " (lat,long) "
position1 = regexpr(",",B$location);    # Character position of ","
position2 = regexpr(")",B$location);    # Character position of ")"

B$Y = as.numeric(substr(B$location, 2, position1-1))              # Y = Latitude
B$X = as.numeric(substr(B$location, position1+1, position2-1))    # X = Longitude

# Erase "callDateTime" "location"     "hour"         "min"          "sec"          "ampm"     
# keep = c("year","month","day","time","Y","X","priority")

```

Dates are not numbered from 1-[highest unique number]
```{r}
keep = c("t","X","Y","priority")
B = B[ , (names(B) %in% keep)]
B = na.omit(B)
head(B)
```
My strategy for converting dates:
df <- transform(df,Date_adj = as.numeric(factor(Date)))


##Check the units for Xbin and Ybin
Do the values have the order of 1,2,3....?
```{r, warning = FALSE}
h = 3    # h IQRs
y1 = quantile(B$Y,.25) - h*IQR(B$Y)
y2 = quantile(B$Y,.75) + h*IQR(B$Y)
x1 = quantile(B$X,.25) - h*IQR(B$X)
x2 = quantile(B$X,.75) + h*IQR(B$X)

Z = as.numeric(B$Y > y1 & B$Y < y2 & B$X > x1 & B$X < x2)

B = B[Z==1,]

B$level = 1*( B$priority=="Non-Emergency") + 3*( B$priority=="Low") + 4*( B$priority=="Medium") + 2*( B$priority=="High") + 6*( B$priority=="Emergency") 
# 1 = black, 2 = red, 3 = green, 4 = blue, 6 = maroon

plot(B$X, B$Y,xlim=c(x1,x2),ylim=c(y1,y2),col=B$level, pch=20)

# Baltimore center
xx = c(-76.67,-76.57); yy = c(39.27,39.32);

plot(B$X, B$Y,xlim=xx,ylim=yy)
points(B$X, B$Y,xlim=xx,ylim=yy,col=B$level, pch=20)
lines(xx,c(1,1)*yy[1] ,lwd=3); lines(xx,c(1,1)*yy[2] ,lwd=3); lines(xx[1]*c(1,1),yy ,lwd=3); lines(xx[2]*c(1,1),yy ,lwd=3);

# Grid
Nbins = 30; DeltaX = (xx[2]-xx[1])/(Nbins +1); DeltaY = (yy[2]-yy[1])/(Nbins +1);

for (k in 0 : (Nbins )){ lines(xx,c(1,1)*(yy[1]+k*DeltaY)); lines(c(1,1)*(xx[1]+k*DeltaX),yy); }
lines(xx,c(1,1)*yy[1] ,lwd=3); lines(xx,c(1,1)*yy[2] ,lwd=3); lines(xx[1]*c(1,1),yy ,lwd=3); lines(xx[2]*c(1,1),yy ,lwd=3);
B$Y
```

```{r}
# Find standardized locations Xbin and Ybin 
Xbin = ceiling((B$X - xx[1])/DeltaX);
Ybin = ceiling((B$Y - yy[1])/DeltaY);
Tbin = B$t - min(B$t) + 1;            # standardized time
Ndays   = max(Tbin)
Xbin
```

```{r}
min(Xbin)
max(Xbin)
```